### 2022-06-21 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@1b4ddd7d8db0b6e031f2dc59c93240bedff693b2 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@16e453a1bff6288438ff77acaba93a8c85556ae7 🚀

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily The coevolutionary mosaic of
\emph{Betacoronavirus} spillover risk from bat hosts}
\vskip 5em

@@

Disease emergence is complex, and is driven not only by animal-human
contact, but also by the underlying evolutionary dynamics in viral
reservoirs (Plowright et al. 2017). Although host richness is often used
as a superficial proxy for spillover risk (Anthony et al. 2017,
Ruiz-Aravena et al. 2022), these approaches oversimplify the relevant
interspecific heterogeneity in immunology, behavior, and other traits,
and therefore overlook unique host pools that allow for the rapid
evolution of highly divergent viruses (Agosta et al. 2010). In the case
of generalist pathogens like betacoronaviruses, there is conceptual and
empirical support to the idea that these community-level mechanisms are
even more important (Power and Mitchell 2004), particularly given that
cross-species transmission may, as a rule, structure viral evolution
more than co-divergence with hosts {[}Geoghegan{]}. This creates a
disconnect between coevolutionary theory (including empirical evidence
from virology) and most existing ecological frameworks for mapping
spillover risk.

The Geographic Mosaic Theory of Coevolution (GMTC; Thompson 2005)
attempts to explicitly connect microevolutionary dynamics to the
macroecology and biogeography of symbiotic interactions. The GMTC posits
that coevolutionary processes among pairs (Thompson 1994) or complexes
(Janzen 1980) of species are structured in space by the rippling effects
of abiotic conditions onto evolutionary mechanism, giving rise to
fragmented systems with different structure and ecologically dynamics
over large spatial extents (see \emph{e.g.} Price 2002). The GMTC
predicts a spatial fragmentation of coevolutionary dynamics under the
joint action of three processes (see notably Gomulkiewicz et al. 2007):
coevolutionary hot- and coldspots, which appear when the intensity of
\emph{interaction} (in terms of reciprocal fitness consequences) varies
spatially; selection mosaics, wherein the intensity of \emph{selection}
varies across space, driven by both the biotic complexity of the
community (locally diverse hosts and viruses are more biotically
complex) and the local favorability of the environment
(\textbf{Thrall2007?}); and trait remixing, which occurs when
coevolutionary dynamics are driven by by the arrival (or departure) of
\emph{functional traits}, through changes in community composition due
to invasions, meta-community dynamics, and disperal.

Here, we apply the GMTC to explore and explain the global biogeography
of betacoronaviruses, the group that includes SARS-CoV, MERS-CoV, and
SARS-CoV-2. In their bat reservoirs, coronaviruses evolve through a mix
of host jumps, recombination among disparate lineages, and, to a lesser
degree, co-divergence with their hosts {[}Anthony{]}---a mix of
mechanisms that creates a complex and nonlinear relationship between
host diversity and viral emergence. Working from a recently published
database of bat hosts of betacoronaviruses, we develop the first global
maps of both host and virus evolutionary distinctiveness and
biogeographic regions for this system. Aiming to explain these patterns,
we develop a generalized framework for applying the GMTC to host-virus
interactions, with a specific emphasis on the potential to create
independent coevolutionary dynamics (and therefore spatial fragmentation
in risk) through heterogeneity. We develop a trivariate risk assessment
system that connects each GMTC mechanism to a quantifiable aspect of
host-virus interactions: (i) viral sharing rates in host communities,
representing the strength of potential interaction between viruses and
any one host (i.e., places where viruses undergo constant host switching
may be coevolutionary coldspots); (ii) the phylogenetic diversity of
hosts, as a proxy for variation in the immunological mechanisms that
antagonize viruses (i.e., the selection mosaic); and (iii) the local
uniqueness of the bat community, representing the potential for viruses
to be exposed to novel host traits (e.g., variation in receptor
sequences). Together, we argue that these can be used to identify and
map the evolutionary drivers that---in conjunction with transmission
processes (e.g., viral prevalence in reservoirs and animal-human contact
rates)--- determine disease emergence risk.

@@

Bats, the second most diverse groups of mammals, are found worldwide and
serve as the main animal reservoir for different strains of
betacoronaviruses (Drexler et al. 2014). This has attracted attention to
areas where high diversity of bats, and therefore presumably high
diversity of betacoronaviruses, can be an important issue for human
health (Calisher et al. 2006, Moratelli and Calisher 2015). By
overlaying the IUCN range maps for confirmed bat hosts of
betacoronaviruses {[}fig.~\ref{fig:richness}; top{]}, we see that the
the main hotspots (both in terms of size and higher values) of host
richness are primarily South-Eastern Asia, parts of Southern Europe, and
to a lesser extent parts of Africa in the -25-0 range of latitudes. The
description of host richness is an important first step towards
understanding risk, as previous research (Anthony et al. 2017, Mollentze
and Streicker 2020) states that locally diverse bat communities could
maintain more viruses and hence, a higher probability of having a
pathogen that could represent a risk for human health.

@@

Nevertheless, locally diverse and virus-rich bat communities could
represent an increased risk of spillover under climate change through
the creation of novel interactions (Carlson et al. 2022), and therefore
the diversity of \emph{Betacoronavirus} strains should similarly be
ccounted for. In fig.~\ref{fig:richness} (bottom), we contrast the
evolutionary distinctiveness of bats and viruses -- this reveals a
slightly different portrait than bat richness alone. Chiropterans can be
classified, from a macro-evolutionary standpoint, as Yangochiroptera and
Yinpterochiroptera elsewhere (Teeling et al. 2005, Springer 2013).
Specifically, we would expect that the so-called ``New World'' group of
bats, being more evolutionary distinct, would also have evolutionary
distinct viruses. Indeed fig.~\ref{fig:richness} (bottom) reveals it to
be the case, and this region harbors a distinct bat-betacoronaviruses
complex. This can be explained by the fact that Yangochiroptera,
although not limited to the western hemisphere, contain the highly
diverse adaptive radiation in the Phyllostomidae (Villalobos and Arita
2010), which is restricted to the western hemisphere. By contrast,
South-Eastern Asia has a lot of non-evolutionary distinct bats, who
nevertheless hosted evolutionary-distinct viruses.

It is noteworthy that outside of South America, viral evolutionary
distinctiveness does not accurately track host diversity, with some
areas having over-distinct viruses (eastern China but, oddly, not the
rest of southeast Asia). There are a number of likely explanations.
First, given the richness of bats in southeast Asia, many
betacoronaviruses likely remain to be discovered in this region. Indeed,
global predictions by Becker et al. (2022) highlight that southeast Asia
is a likely hostpot of unconfirmed hosts of betacoronaviruses, which
would likely result in additional viral discoveries. This idea is
unsurprising given the growing realization, especially since the
emergence of SARS-CoV-2, that unique lineages of similar viruses are
widespread in bats but still mostly undescribed. The most distinct
bats-betacoronaviruses complex is found in South America, a region with
a comparatively lower number of hosts; this matches with the isolation
through viariance of the host group, and may highlight a different
co-evolutionary dynamic. Alternatively, this distinctiveness hostpot may
be a product of under-sampling: South-America is one of the places where
the fewest \emph{Betacoronavirus} sequences have been discovered
(Anthony et al. 2017, Olival et al. 2017, Allen et al. 2017), resulting
in sparser phylogenetic tree, thereby artificially inflating
distinctiveness. Adding more viruses would bring the distinctiveness of
known sequences down. Finally, South-America is the range of
Phyllostomidae, a group of bats that underwent explosive diversification
events (Villalobos and Arita 2010), which may drive the emergence of
multiple viral lineages.

@@

Despite the difference in evolutionary distinctiveness globally, there
are reasons to expect that the phylogeography of bats and
betacoronaviruses should show some degree of congruence (Letko et al.
2020, Van Brussel and Holmes 2022). In particular, this should be the
case if viruses can circulate among hosts and co-evolve with local hosts
communities, making their evolutionary process more than a byproduct of
host evolution. High density of hosts sharing the same virus (albeit
possibly different strains) can drive or result from evolution of the
bat antiviral immune system, resulting in spatially distinct
immunological responses, as evidenced in several bat species (Banerjee
et al. 2020). Immune characteristics that allow bats to be better
adapted to infection by emerging viruses (Gorbunova et al. 2020, Irving
et al. 2021), in addition to being hardcoded in their genome (Jebb et
al. 2020), may be related to a wide variety of diets (Banerjee et al.
2020, Moreno Santillán et al. 2021), themselves likely to be driven by
spatial effects, especially at the local scale -- bats, indeed, occupy a
variety of environments, and therefore display a variety of adaptations
to these environments (Muylaert et al. 2022).

@@

In fig.~\ref{fig:biogeo}, we show a projection of the phylogeographic
signal of bats (top) and viruses (bottom) in space; the distinct
groupings (represented by different colors symbolizing positions in the
subspace formed by the first two axes of the PCoA) are essentially
equivalent between the two groups, and can be coarsely delineated as
southeast Asia, Eurasia above a northing of 25, and Africa and south
America. These results suggest that, although the evolutionary
distinctiveness of the bat-betacoronaviruses complex varies spatially,
the system shows an important degree of spatial consistency, with a
reduced number of bioregions. Available information describing the
spillover of zoonotic betacoronaviruses of bat origin where data was
available before and up through the COVID-19 pandemic puts spillover
events of SARS-CoV-2 in Wuhan, China; SARS-CoV in Guangdong, China based
on the presence of closest known viruses circulating in nature, and a
nearby location where serological (antibody) evidence has indicated
human exposure to SARS-like viruses (Wang et al. 2018); MERS-CoV in
Saudi Arabia based on index cases available from a recently-published
compendium of cases (Ramshaw et al. 2019). For the latest event, most if
not all index cases are presumed to be camel-to-human transmission, and
the precise origin point (if it exists) of MERS-CoV in bats is
uncertain. Recent recombinant canine coronavirus spillover events in
Haiti (Lednicky et al. 2021) and Europe (Vlasova et al. 2022) are not
relevant here, as bats' involvement in these cycles of transmission have
been supposed to be non-existent. These index cases fall within
different phylogeographic bioregions (fig.~\ref{fig:biogeo}), which
further highlight the issue that different host-virus sub-systems may
lead to widespread emergence.

@@

As host richness, joint distinctiveness, or phylogeographic structure
suggest that the bat-betacoronaviruses complex is globally fragmented
enough to give rise to both different levels of risk (as evidenced by
the spatial location of spillover events) and different types of
co-evolutionary dynamics, we turn to the Geographic Mosaic Theory of
Coevolution (Thompson 2005) to provide a measure of risk accounting for
multiple processes. In fig.~\ref{fig:trivariate}, we overlapped three
components of spillover risk: viral sharing, \emph{i.e.} the chance that
two bats will share viruses overall; Local Contribution to Beta
Diversity, \emph{i.e.} the fact that a bat community is compositionally
unique compared to the average compositional similarity across the
entire system; finally, host phylogenetic diversity, \emph{i.e.} how
dispersed the bats in a location are within the tree of life. This
approach leads to the definition of broad biogeographic regions of risk,
where the same color represents the same type of risk. By way of
constrat to figures fig.~\ref{fig:richness} and fig.~\ref{fig:biogeo},
these regions do not necessarilly overlap with previous spatial
partitions of the bat-betacoronaviruses complex.

@@

From the perspective of spillover risk, the most important combination
of factors is a high phylogenetic diversity of hosts with low viral
sharing; this, essentially, means that very different betacoronaviruses
could co-exist within the same place. This is particularly the case
given that betacoronaviruses often evolve and even achieve host shifts
through recombination, which requires the co-occurrence of sufficiently
distinct viruses to be a major driver of emergence. In
fig.~\ref{fig:trivariate}, this corresponds to yellow to pale green
areas, which are essentially limited to South-Eastern Asia, and to some
part of Sub-Saharan Africa. Adopting a geographic mosaic theory
perspective on risk, other regions of the world are of lesser concern
fig.~\ref{fig:risk}. Our risk decomposition does not account for viral
diversity or distinctiveness. The simple rationale behind it is that the
acquisition of viral data is rarely disconnected from the acquisition of
host data. There are more sources of information on hosts than on
viruses, allowing to develop a host-centric perspective on risk
(although this estimate would more accurate with viral traits related to
\emph{e.g.} ability to switch hosts or pathogenic potential). Areas with
high bat diversity and high turnover \emph{may} facilitate the
evolutionary radiation of viruses, matching previous findings that the
diversification of bat coronaviruses is driven largely by host shifts
(inter-genus or higher levels of cross-species transmission) and, to a
lesser degree, cospeciation and sharing (intra-genus cross-species
transmission; Anthony et al. 2017). This diversification is not an
actual risk factor for spillover itself, but acts downstream of a
spillover event by increasing the random chance of the emergence of a
virus with the raw genomic components required for the potential to
infect humans.

@@

From another perspective, areas of high host uniqueness and virus
sharing (red-to-pink) could provide hotspots of \emph{Betacoronavirus}
risk through mixing of unique viruses (via codivergence) and in turn
recombination. Under our framework, such a hotspot was identified in
Madagascar, where most bat species are endemic following evolutionary
divergence from sister species in both African and Asian continents
(\emph{e.g.} Shi et al. 2014). Recent surveillance (Kettenburg et al.
2022) has identified a novel \emph{Betacoronavirus} (in the subgenus
\emph{Nobecovirus}) in Madagascar-endemic pteropid bat species
(\emph{Pteropus rufus}, \emph{Rousettus madagascariensis}), emphasizing
strong proof of principle in model predictions.

@@

Based on the previous result, we extracted the risk component from the
composite map (see Methods), to provide a single measure of risk varying
between 0 and 1. This measure is presented in fig.~\ref{fig:risk}. As
this map represents the potential risk, it must be weighed by the
potential for contacts with humans. As a proxy for this measure, we used
the proportion of build/urban land from the EarthEnv dataset: this is a
reasonable proxy for the density of humans per unit area, which
increases the probability of pathogen spread more widely (Hazarie et al.
2021). Since human activity is required to amplify the frequency of
virus encounters and thus create areas of viral amplification, mapping
the potential risk against measures of land use is required to generate
a more actionable assessment of risk. This map is presented in
fig.~\ref{fig:compound}. Most of South America and Europe are at
comparatively lower risk, as although densely populated, settlements
tend to be in areas with lower potential risk. Regions like Malaysia and
the North coast of Australia have a high risk component, but should
represent a relatively lower effective risk due to low human density.
However, this mapping reveals that South-East Asia, the Indian
subcontinent, and parts of sub-Saharan Africa, are at high risk due to
the overlap between built areas and bat communities representing more
opportunities for cross-species transmission of betacoronaviruses. In
looking for the origins of SARS in China, Xu et al. (2004) present
serological evidence that strongest human-animal contact results in
higher risk of virus exposure, regardless of the animal species, but
that different types of contact had different impacts. Ideally,
finer-grained information about human activity (rather than human
presence through anthropisation) could allow to partition this risk
further, alebit at the cost of more hypotheses required to estimate the
amount of risk represented by each activity. Our map of risk overlays
with recent results from Cohen et al. (2022) -- areas of purported high
risk/diversitifcation potential (Madagascar, South-America) overlay with
sampling gaps for \emph{Betacoronavirus}, stressing the need for
spatially targeted monitoring and discovery.

@@

Bats are important reservoir hosts for different classes of
microorganisms, many of which a threat to human health (Letko et al.
2020, Van Brussel and Holmes 2022). Chiropterans emerged around 64
million years ago and are one of the most diverse mammalian orders, with
an estimated richness of more than 1400 species (Peixoto et al. 2018,
Simmons and Cirranello 2020). They exhibit a broad variety of habitat
use, behaviour, and feeding strategies, putting them at key positions in
the delivery and provisioning of several ecosystem services, tied to
important ecosystem-derived benefits to human (Kasso and Balakrishnan
2013). For example, bats are an essential component of many
seed-dispersal networks (Mello et al. 2011). Over two-thirds of bats are
know to be either obligate or facultative insectivores, therefore
actively contributing for agricultural pest control (Williams-Guillén et
al. 2008, Voigt and Kingston 2016), and vectors of pathogens that put a
risk on human health (Gonsalves et al. 2013a, b). Because bats are
globally distributed and have a long evolutionary history,
phylogeographic and biogeographic approaches are required to shed light
on the contemporary distribution of coevolutionary processes between
bats and the pathogens they host. Not all areas in which bats, viruses,
and human are co-occuring are facing a risk of spillover towards human
populations, and the areas in which this risk exist may not be facing
risks of the same nature and magnitude.

>>>>

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily The coevolutionary mosaic of betacoronavirus
emergence risk}
\vskip 5em

@@

Disease emergence is complex, and is driven not only by animal-human
contact, but also by the underlying evolutionary dynamics in viral
reservoirs.\textsuperscript{1} Although host richness is often used as a
superficial proxy for spillover risk,\textsuperscript{2,3} these
approaches oversimplify the relevant interspecific heterogeneity in
immunology, behavior, and other traits, and therefore overlook unique
host pools that allow for the rapid evolution of highly divergent
viruses.\textsuperscript{4} In the case of generalist pathogens like
betacoronaviruses, there is conceptual and empirical support to the idea
that these community-level mechanisms are even more
important,\textsuperscript{5} particularly given that cross-species
transmission may, as a rule, structure viral evolution more than
co-divergence with hosts {[}Geoghegan{]}. This creates a disconnect
between coevolutionary theory (including empirical evidence from
virology) and most existing ecological frameworks for mapping spillover
risk.

The Geographic Mosaic Theory of Coevolution\textsuperscript{GMTC; 6}
attempts to explicitly connect microevolutionary dynamics to the
macroecology and biogeography of symbiotic interactions. The GMTC posits
that coevolutionary processes among pairs\textsuperscript{7} or
complexes\textsuperscript{8} of species are structured in space by the
rippling effects of abiotic conditions onto evolutionary mechanism,
giving rise to fragmented systems with different structure and
ecologically dynamics over large spatial extents.\textsuperscript{see
\emph{e.g.} 9} The GMTC predicts a spatial fragmentation of
coevolutionary dynamics under the joint action of three
processes:\textsuperscript{see notably 10} coevolutionary hot- and
coldspots, which appear when the intensity of \emph{interaction} (in
terms of reciprocal fitness consequences) varies spatially; selection
mosaics, wherein the intensity of \emph{selection} varies across space,
driven by both the biotic complexity of the community (locally diverse
hosts and viruses are more biotically complex) and the local
favorability of the environment;\textsuperscript{11} and trait remixing,
which occurs when coevolutionary dynamics are driven by by the arrival
(or departure) of \emph{functional traits}, through changes in community
composition due to invasions, meta-community dynamics, and disperal.

Here, we apply the GMTC to explore and explain the global biogeography
of betacoronaviruses, the group that includes SARS-CoV, MERS-CoV, and
SARS-CoV-2. In their bat reservoirs, coronaviruses evolve through a mix
of host jumps, recombination among disparate lineages, and, to a lesser
degree, co-divergence with their hosts---\textsuperscript{2}a mix of
mechanisms that creates a complex and nonlinear relationship between
host diversity and viral emergence. Working from a recently published
database of bat hosts of betacoronaviruses, we develop the first global
maps of both host and virus evolutionary distinctiveness and
biogeographic regions for this system. Aiming to explain these patterns,
we develop a generalized framework for applying the GMTC to host-virus
interactions, with a specific emphasis on the potential to create
independent coevolutionary dynamics (and therefore spatial fragmentation
in risk) through heterogeneity. We develop a trivariate risk assessment
system that connects each GMTC mechanism to a quantifiable aspect of
host-virus interactions: (i) viral sharing rates in host communities,
representing the strength of potential interaction between viruses and
any one host (i.e., places where viruses undergo constant host switching
may be coevolutionary coldspots); (ii) the phylogenetic diversity of
hosts, as a proxy for variation in the immunological mechanisms that
antagonize viruses (i.e., the selection mosaic); and (iii) the local
uniqueness of the bat community, representing the potential for viruses
to be exposed to novel host traits (e.g., variation in receptor
sequences). Together, we argue that these can be used to identify and
map the evolutionary drivers that---in conjunction with transmission
processes (e.g., viral prevalence in reservoirs and animal-human contact
rates)--- determine disease emergence risk.

@@

Bats, the second most diverse groups of mammals, are found worldwide and
serve as the main animal reservoir for different strains of
betacoronaviruses.\textsuperscript{12} This has attracted attention to
areas where high diversity of bats, and therefore presumably high
diversity of betacoronaviruses, can be an important issue for human
health.\textsuperscript{13,14} By overlaying the IUCN range maps for
confirmed bat hosts of betacoronaviruses {[}fig.~\ref{fig:richness};
top{]}, we see that the the main hotspots (both in terms of size and
higher values) of host richness are primarily South-Eastern Asia, parts
of Southern Europe, and to a lesser extent parts of Africa in the -25-0
range of latitudes. The description of host richness is an important
first step towards understanding risk, as previous
research\textsuperscript{2,15} states that locally diverse bat
communities could maintain more viruses and hence, a higher probability
of having a pathogen that could represent a risk for human health.

@@

Nevertheless, locally diverse and virus-rich bat communities could
represent an increased risk of spillover under climate change through
the creation of novel interactions,\textsuperscript{16} and therefore
the diversity of \emph{Betacoronavirus} strains should similarly be
ccounted for. In fig.~\ref{fig:richness} (bottom), we contrast the
evolutionary distinctiveness of bats and viruses -- this reveals a
slightly different portrait than bat richness alone. Chiropterans can be
classified, from a macro-evolutionary standpoint, as Yangochiroptera and
Yinpterochiroptera elsewhere.\textsuperscript{17,18} Specifically, we
would expect that the so-called ``New World'' group of bats, being more
evolutionary distinct, would also have evolutionary distinct viruses.
Indeed fig.~\ref{fig:richness} (bottom) reveals it to be the case, and
this region harbors a distinct bat-betacoronaviruses complex. This can
be explained by the fact that Yangochiroptera, although not limited to
the western hemisphere, contain the highly diverse adaptive radiation in
the Phyllostomidae,\textsuperscript{19} which is restricted to the
western hemisphere. By contrast, South-Eastern Asia has a lot of
non-evolutionary distinct bats, who nevertheless hosted
evolutionary-distinct viruses.

It is noteworthy that outside of South America, viral evolutionary
distinctiveness does not accurately track host diversity, with some
areas having over-distinct viruses (eastern China but, oddly, not the
rest of southeast Asia). There are a number of likely explanations.
First, given the richness of bats in southeast Asia, many
betacoronaviruses likely remain to be discovered in this region. Indeed,
global predictions by\textsuperscript{20} highlight that southeast Asia
is a likely hostpot of unconfirmed hosts of betacoronaviruses, which
would likely result in additional viral discoveries. This idea is
unsurprising given the growing realization, especially since the
emergence of SARS-CoV-2, that unique lineages of similar viruses are
widespread in bats but still mostly undescribed. The most distinct
bats-betacoronaviruses complex is found in South America, a region with
a comparatively lower number of hosts; this matches with the isolation
through viariance of the host group, and may highlight a different
co-evolutionary dynamic. Alternatively, this distinctiveness hostpot may
be a product of under-sampling: South-America is one of the places where
the fewest \emph{Betacoronavirus} sequences have been
discovered,\textsuperscript{2,21,22} resulting in sparser phylogenetic
tree, thereby artificially inflating distinctiveness. Adding more
viruses would bring the distinctiveness of known sequences down.
Finally, South-America is the range of Phyllostomidae, a group of bats
that underwent explosive diversification events,\textsuperscript{19}
which may drive the emergence of multiple viral lineages.

@@

Despite the difference in evolutionary distinctiveness globally, there
are reasons to expect that the phylogeography of bats and
betacoronaviruses should show some degree of
congruence.\textsuperscript{23,24} In particular, this should be the
case if viruses can circulate among hosts and co-evolve with local hosts
communities, making their evolutionary process more than a byproduct of
host evolution. High density of hosts sharing the same virus (albeit
possibly different strains) can drive or result from evolution of the
bat antiviral immune system, resulting in spatially distinct
immunological responses, as evidenced in several bat
species.\textsuperscript{25} Immune characteristics that allow bats to
be better adapted to infection by emerging
viruses,\textsuperscript{26,27} in addition to being hardcoded in their
genome,\textsuperscript{28} may be related to a wide variety of
diets,\textsuperscript{25,29} themselves likely to be driven by spatial
effects, especially at the local scale -- bats, indeed, occupy a variety
of environments, and therefore display a variety of adaptations to these
environments.\textsuperscript{30}

@@

In fig.~\ref{fig:biogeo}, we show a projection of the phylogeographic
signal of bats (top) and viruses (bottom) in space; the distinct
groupings (represented by different colors symbolizing positions in the
subspace formed by the first two axes of the PCoA) are essentially
equivalent between the two groups, and can be coarsely delineated as
southeast Asia, Eurasia above a northing of 25, and Africa and south
America. These results suggest that, although the evolutionary
distinctiveness of the bat-betacoronaviruses complex varies spatially,
the system shows an important degree of spatial consistency, with a
reduced number of bioregions. Available information describing the
spillover of zoonotic betacoronaviruses of bat origin where data was
available before and up through the COVID-19 pandemic puts spillover
events of SARS-CoV-2 in Wuhan, China; SARS-CoV in Guangdong, China based
on the presence of closest known viruses circulating in nature, and a
nearby location where serological (antibody) evidence has indicated
human exposure to SARS-like viruses;\textsuperscript{31} MERS-CoV in
Saudi Arabia based on index cases available from a recently-published
compendium of cases.\textsuperscript{32} For the latest event, most if
not all index cases are presumed to be camel-to-human transmission, and
the precise origin point (if it exists) of MERS-CoV in bats is
uncertain. Recent recombinant canine coronavirus spillover events in
Haiti\textsuperscript{33} and Europe\textsuperscript{34} are not
relevant here, as bats' involvement in these cycles of transmission have
been supposed to be non-existent. These index cases fall within
different phylogeographic bioregions (fig.~\ref{fig:biogeo}), which
further highlight the issue that different host-virus sub-systems may
lead to widespread emergence.

@@

As host richness, joint distinctiveness, or phylogeographic structure
suggest that the bat-betacoronaviruses complex is globally fragmented
enough to give rise to both different levels of risk (as evidenced by
the spatial location of spillover events) and different types of
co-evolutionary dynamics, we turn to the Geographic Mosaic Theory of
Coevolution\textsuperscript{6} to provide a measure of risk accounting
for multiple processes. In fig.~\ref{fig:trivariate}, we overlapped
three components of spillover risk: viral sharing, \emph{i.e.} the
chance that two bats will share viruses overall; Local Contribution to
Beta Diversity, \emph{i.e.} the fact that a bat community is
compositionally unique compared to the average compositional similarity
across the entire system; finally, host phylogenetic diversity,
\emph{i.e.} how dispersed the bats in a location are within the tree of
life. This approach leads to the definition of broad biogeographic
regions of risk, where the same color represents the same type of risk.
By way of constrat to figures fig.~\ref{fig:richness} and
fig.~\ref{fig:biogeo}, these regions do not necessarilly overlap with
previous spatial partitions of the bat-betacoronaviruses complex.

@@

From the perspective of spillover risk, the most important combination
of factors is a high phylogenetic diversity of hosts with low viral
sharing; this, essentially, means that very different betacoronaviruses
could co-exist within the same place. This is particularly the case
given that betacoronaviruses often evolve and even achieve host shifts
through recombination, which requires the co-occurrence of sufficiently
distinct viruses to be a major driver of emergence. In
fig.~\ref{fig:trivariate}, this corresponds to yellow to pale green
areas, which are essentially limited to South-Eastern Asia, and to some
part of Sub-Saharan Africa. Adopting a geographic mosaic theory
perspective on risk, other regions of the world are of lesser concern
fig.~\ref{fig:risk}. Our risk decomposition does not account for viral
diversity or distinctiveness. The simple rationale behind it is that the
acquisition of viral data is rarely disconnected from the acquisition of
host data. There are more sources of information on hosts than on
viruses, allowing to develop a host-centric perspective on risk
(although this estimate would more accurate with viral traits related to
\emph{e.g.} ability to switch hosts or pathogenic potential). Areas with
high bat diversity and high turnover \emph{may} facilitate the
evolutionary radiation of viruses, matching previous findings that the
diversification of bat coronaviruses is driven largely by host shifts
(inter-genus or higher levels of cross-species transmission) and, to a
lesser degree, cospeciation and sharing.\textsuperscript{intra-genus
cross-species transmission; 2} This diversification is not an actual
risk factor for spillover itself, but acts downstream of a spillover
event by increasing the random chance of the emergence of a virus with
the raw genomic components required for the potential to infect humans.

@@

From another perspective, areas of high host uniqueness and virus
sharing (red-to-pink) could provide hotspots of \emph{Betacoronavirus}
risk through mixing of unique viruses (via codivergence) and in turn
recombination. Under our framework, such a hotspot was identified in
Madagascar, where most bat species are endemic following evolutionary
divergence from sister species in both African and Asian
continents.\textsuperscript{\emph{e.g.} 35} Recent
surveillance\textsuperscript{36} has identified a novel
\emph{Betacoronavirus} (in the subgenus \emph{Nobecovirus}) in
Madagascar-endemic pteropid bat species (\emph{Pteropus rufus},
\emph{Rousettus madagascariensis}), emphasizing strong proof of
principle in model predictions.

@@

Based on the previous result, we extracted the risk component from the
composite map (see Methods), to provide a single measure of risk varying
between 0 and 1. This measure is presented in fig.~\ref{fig:risk}. As
this map represents the potential risk, it must be weighed by the
potential for contacts with humans. As a proxy for this measure, we used
the proportion of build/urban land from the EarthEnv dataset: this is a
reasonable proxy for the density of humans per unit area, which
increases the probability of pathogen spread more
widely.\textsuperscript{37} Since human activity is required to amplify
the frequency of virus encounters and thus create areas of viral
amplification, mapping the potential risk against measures of land use
is required to generate a more actionable assessment of risk. This map
is presented in fig.~\ref{fig:compound}. Most of South America and
Europe are at comparatively lower risk, as although densely populated,
settlements tend to be in areas with lower potential risk. Regions like
Malaysia and the North coast of Australia have a high risk component,
but should represent a relatively lower effective risk due to low human
density. However, this mapping reveals that South-East Asia, the Indian
subcontinent, and parts of sub-Saharan Africa, are at high risk due to
the overlap between built areas and bat communities representing more
opportunities for cross-species transmission of betacoronaviruses. In
looking for the origins of SARS in China,\textsuperscript{38} present
serological evidence that strongest human-animal contact results in
higher risk of virus exposure, regardless of the animal species, but
that different types of contact had different impacts. Ideally,
finer-grained information about human activity (rather than human
presence through anthropisation) could allow to partition this risk
further, alebit at the cost of more hypotheses required to estimate the
amount of risk represented by each activity. Our map of risk overlays
with recent results from\textsuperscript{39} -- areas of purported high
risk/diversitifcation potential (Madagascar, South-America) overlay with
sampling gaps for \emph{Betacoronavirus}, stressing the need for
spatially targeted monitoring and discovery.

@@

Bats are important reservoir hosts for different classes of
microorganisms, many of which a threat to human
health.\textsuperscript{23,24} Chiropterans emerged around 64 million
years ago and are one of the most diverse mammalian orders, with an
estimated richness of more than 1400 species.\textsuperscript{40,41}
They exhibit a broad variety of habitat use, behaviour, and feeding
strategies, putting them at key positions in the delivery and
provisioning of several ecosystem services, tied to important
ecosystem-derived benefits to human.\textsuperscript{42} For example,
bats are an essential component of many seed-dispersal
networks.\textsuperscript{43} Over two-thirds of bats are know to be
either obligate or facultative insectivores, therefore actively
contributing for agricultural pest control,\textsuperscript{44,45} and
vectors of pathogens that put a risk on human
health.\textsuperscript{46,47} Because bats are globally distributed and
have a long evolutionary history, phylogeographic and biogeographic
approaches are required to shed light on the contemporary distribution
of coevolutionary processes between bats and the pathogens they host.
Not all areas in which bats, viruses, and human are co-occuring are
facing a risk of spillover towards human populations, and the areas in
which this risk exist may not be facing risks of the same nature and
magnitude.

Here, we propose a simple freamework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the previously-neglected adaptive radiation of
sarbecoviruses outside of southern China and throughout southeast Asia.
In doing so, it advances ecological theory beyond the current state of
the art for global maps of emergence risk. For example, previous studies
that have used host richness as proxy have predicted a high diversity of
unsampled bat viruses,\textsuperscript{22} bat
coronaviruses,\textsuperscript{2} and even specifically
betacoronaviruses\textsuperscript{20} in both the Amazon and southeast
Asia. While we find that both regions are characterized by highly
divergent host and viral communities, our framework identifies key
differences between the regions. We find that Latin America is a hotspot
of both host and viral distinctiveness, suggesting that this branch of
the bat-betacoronavirus complex may be undergoing independent
evolutionary dynamics from the rest of the global pool, but with limited
potential for viral diversification--- a finding that is supported by
previous work indicating a higher rate of codivergence in Latin
America.\textsuperscript{2} In contrast, in southeast Asia, host
richness and viral distinctiveness are high but sharing is low; this
suggests a different type of evolutionary dynamics that could generate
high local diversity of viruses through host switching and viral
recombination (see \emph{e.g.},\textsuperscript{48} as well as the
discovery of recombinant viruses that share genetic material from both
the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
lineage).\textsuperscript{49} Both of these regions are priority areas
for sampling, especially given predictions that they contain many bat
hosts of undiscovered betacoronaviruses.\textsuperscript{20,39} However,
both the evolutionary and ecological aspects of emergence risk are
likely higher in southeast Asia---a fact that will only become more
relevant, as bats track shifting climates and exchange viruses with
other species, creating a hotspot of cross-species transmission unique
to the region.\textsuperscript{16}

{[}Tim tinker with para about conservation{]} There are several factors
that drive changes in the diversity of bats,\textsuperscript{50} but
human activities' effects on the ecosystem (like modifications of land
use) could significantly decrease it. Therefore, it can be suggested
that changes in the diversity of betacoronaviruses in bats are linked to
their biogeographic variation, and human population density and other
anthropogenic factors are decisive moderators for its implications in
public health. With the increase of contact between humans and potential
hosts, we also increase the risk of emergence of novel
diseases,\textsuperscript{51} as previous studies on RNA viruses suggest
the importance of host phylogeography at the time of virus
dispersal.\textsuperscript{52} One of these scenarios where interaction
between bats and humans can occur can be seed dispersal in tropical
agroecosystems. It opens the discussion of whether the fruits thrown by
bats not only disperse seeds but could also be a source of indirect
interaction between viruses of bat origin and
humans.\textsuperscript{53} This represents a challenge for conservation
strategies and disease ecology since some areas can haveboth potential
for the acquisition of zoonotic viruses and bat-human interactions; in
particular, the challenge lies in the fact that actual exposure must
then be quantified accounting for several transmission scenarios,
including both direct and indirect bat - human interaction.
